---
title: "Explore Acoustic Trawl Data"
format: html
editor: source
---

```{r}
# explore data
library(tidyverse)
library(here)
library(readxl)
library(viridis)
library(sf)
library(rnaturalearth)
library(magick)

# plot theme
pt <- theme_minimal()+theme(panel.border=element_rect(color="black",fill=NA))
theme_set(pt)

# backgroundf map
# spatial background map
# load west cost land for mapping
coast <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>%
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")
coastcn <- ne_countries(country="Canada",scale=50,returnclass='sf')%>%
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")

coast <- st_union(coast,coastcn)
```

```{r}
source(here('code','helper_fxns.R'))
```

# Import Data

There are a lot of files in each folder. Let's try to bind together the un-kriged, aged output for all years into a useable format.

```{r}
# general file name
fn <- "EchoPro_un-kriged_aged_output"
# all files in the FEAT data fitting that match the string above
fls <- list.files(here('data','CONFIDENTIAL','FEAT'),recursive = T,full.names = T) %>% str_subset(fn)
# a couple of manual removals
fls <- fls[-c(23:44,47:50,53:56,61:66)]

# in the aged output, sheet 1 is sexes combined, sheet 2 is males, sheet 3 is females
dat1 <- readxl::read_xlsx(fls[[1]],sheet = 2,skip = 1)
dat2 <- readxl::read_xlsx(fls[[1]],sheet=3,skip=1)
dat <- bind_rows(dat1,dat2)

import_aged_year <- function(yr){
  to_import <- fls %>% str_subset(paste(yr))
  males <- map_df(to_import,function(fl) {
    read_xlsx(fl,sheet = 2,skip = 1) %>% 
    mutate(sex="male") %>% 
    rename(wgt=wgt_male)
  })
  females <- map_df(to_import,function(fl) {
    read_xlsx(fl,sheet = 3,skip = 1) %>% 
    mutate(sex="female") %>% 
    rename(wgt=wgt_female)
  })
  bind_rows(males,females) %>% mutate(year=yr)
}

```

```{r}
dat_aged <- map_df(c(1995,1998,seq(2001,2023,by=2),2012),import_aged_year)
# save for later
write_rds(dat_aged,here('data','CONFIDENTIAL','un-kriged_aged_output_allyears.rds'))
```

# Age Patterns

Bubble plot of age structure?

```{r}
dat_agesex_comps_year <- dat_aged %>% 
  group_by(year,sex) %>% 
  summarise(across(`1`:`20`,\(x) sum(x,na.rm=T))) %>% 
  ungroup()

dat_agesex_long <- dat_agesex_comps_year %>% 
  pivot_longer(`1`:`20`,values_to = "n",names_to="age") %>% 
  mutate(age=as.integer(age)) %>% 
  # add age props
  group_by(year,sex) %>% 
  mutate(ageprop=n/sum(n,na.rm=T)) %>% 
  ungroup()

p_agesex_comps_year <- dat_agesex_long %>%
  ggplot(aes(year,age,size=n))+
  facet_wrap(~sex)+
  geom_point()
p_agesex_comps_year

dat_agesex_long %>% 
  pivot_wider(names_from=sex,values_from = n) %>% 
  ggplot(aes(female,male))+geom_point()+
  geom_smooth()+
  geom_abline(slope=1,intercept=0,linetype=2)

# as proportions
p_agesex_props_year <- dat_agesex_long %>%
  ggplot(aes(year,age,size=ageprop))+
  geom_point()+
  facet_wrap(~sex)
p_agesex_props_year

```

# Spatial

```{r}
dat_aged_sf <- dat_aged %>% 
  mutate(nonzero=wgt>0) %>% 
  st_as_sf(coords=c("Lon","Lat"),crs=4326)%>%
  st_transform(crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")
```

```{r bbox}
buff <- st_bbox(dat_aged_sf)
coast <- coast %>% st_crop(buff)
```

```{r,fig.height=12,fig.width=8}
# ggplot()+
#   geom_sf(data=coast,fill='gray80')+
#   geom_sf(data=dat_aged_sf)+
#   facet_wrap(~year)

# plot and save maps for each year
ggplot()+
  geom_sf(data=coast,fill='gray80')+
  geom_sf(data=dat_aged_sf %>% filter(year==2023,sex=='female'),size=0.5,aes(col=nonzero))+
  labs(color="Hake Presence")+
  coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
  scale_color_manual(values=c("gray80","red"))+
  theme(legend.position=c(0.7,0.8),legend.background = element_rect(color="white"))

walk(unique(dat_aged_sf$year),function(yr){
  p <- ggplot()+
    geom_sf(data=coast,fill='gray80')+
    geom_sf(data=dat_aged_sf %>% filter(year==yr,sex=='female'),size=0.5,aes(col=nonzero))+
    labs(color="Hake Presence")+
    coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
    scale_color_manual(values=c("gray80","red"))+
    theme(legend.position=c(0.7,0.8),legend.background = element_rect(color="white"))
  ggsave(here('model output','maps',paste0('FEAT samples map ',yr,'.png')),p,w=8,h=12)
})
```

Rotated and facetted map of samples

```{r}
coast_rot <- coast %>% 
  st_transform(st_crs(dat_aged_sf)) %>% 
  rotate_sf(.,pi/9) %>%
  st_set_crs(st_crs(dat_aged_sf))

dat_aged_rot <- dat_aged_sf %>% 
  filter(sex=='female') %>% 
  rotate_sf(.,pi/9) %>%
  st_set_crs(st_crs(dat_aged_sf))
bbox_rot <- st_bbox(dat_aged_rot)

p_samps_all_yrs <- ggplot()+
  geom_sf(data=coast_rot,fill='gray80')+
  geom_sf(data=dat_aged_rot,size=0.2,aes(col=nonzero))+
  labs(color="Hake Presence")+
  coord_sf(xlim = c(bbox_rot[1],bbox_rot[3]),ylim=c(bbox_rot[2],bbox_rot[4]),datum=NA)+
  # coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
  scale_color_manual(values=c("gray80","red"))+
  facet_wrap(~year,nrow=2)+
  theme(legend.background = element_rect(color="white"))

ggsave(here('model output','maps',paste0('FEAT samples map all years.png')),p_samps_all_yrs,w=10,h=8)
```


## Spatial age props

```{r,fig.height=12,fig.width=8}
dat_spatial_ageprop <- dat_aged_sf %>% 
  filter(wgt>0) %>% 
  mutate(sampnum=row_number()) %>% 
  pivot_longer(`1`:`20`,values_to = "n",names_to="age") %>% 
  mutate(age=as.integer(age)) %>% 
  # add age props
  group_by(year,sampnum) %>% 
  mutate(ageprop=n/sum(n,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(ageprop=na_if(ageprop,0))

p <- ggplot()+
  geom_sf(data=coast,fill='gray80')+
  geom_sf(data=dat_aged_sf %>% filter(year==2003),col='gray90',size=0.25)+
  geom_sf(data= dat_spatial_ageprop %>% filter(year==2003,sex=='female',age==20),size=0.5,aes(col=ageprop))+
  labs(color="Proportion")+
  coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
  scale_color_viridis(direction = -1,limits=c(0,1),na.value='gray80')+
  theme(legend.position=c(0.7,0.8),legend.background = element_rect(color="white"))
p
yrs_ages <- crossing(year=unique(dat_aged$year),age=2:20) %>% arrange(age,year)

walk2(yrs_ages$year,yrs_ages$age,function(yr,a){
  d1 <- dat_spatial_ageprop %>% filter(year==yr,sex=='female',age==a)
  p <- ggplot()+
    geom_sf(data=coast,fill='gray80',size=0.25)+
    geom_sf(data=dat_aged_sf %>% filter(year==yr),col='gray90',size=0.5)+
    geom_sf(data=d1,size=0.5,aes(col=ageprop))+
    labs(color="Proportion")+
    coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
    scale_color_viridis(direction = -1,limits=c(0,1))+
    labs(title=paste0(yr,": Age ",a))+
    theme(legend.position=c(0.7,0.8),legend.background = element_rect(color="white"))
  ggsave(here('model output','maps',paste0("age ",a),paste0('Map ',yr,' Age ',a,' proportions.png')),p,w=8,h=12)
})
```

Spatial age props, binned by "juvenile" (ages 2-3) and "adult" (ages 4+)

```{r,fig.height=12,fig.width=8}
samp_info <- dat_spatial_ageprop %>% mutate(x=st_coordinates(.)[,1],y=st_coordinates(.)[,2]) %>% st_set_geometry(NULL) %>%
  distinct(sampnum,x,y)
dat_spatial_ageprop_binned <- dat_spatial_ageprop %>% 
  st_set_geometry(NULL) %>% 
  mutate(binclass=ifelse(age%in%c(1,2,3),"juvenile","adult")) %>% 
  group_by(year,sampnum,binclass) %>% 
  summarise(totn=sum(n,na.rm=T)) %>% 
  group_by(year,sampnum) %>% 
  mutate(prop=totn/sum(totn)) %>% 
  ungroup() %>% 
  mutate(prop=na_if(prop,0)) %>% 
  left_join(samp_info,by=join_by(sampnum)) %>%
  st_as_sf(coords=c('x','y'),crs = "+proj=utm +zone=10 +datum=WGS84 +units=km")
  
yrs_stages <- crossing(year=unique(dat_aged$year),stage=c("juvenile","adult")) %>% arrange(stage,year)

walk2(yrs_stages$year,yrs_stages$stage,function(yr,stg){
  d1 <- dat_spatial_ageprop_binned %>% filter(year==yr,binclass==stg)
  p <- ggplot()+
    geom_sf(data=coast,fill='gray80',size=0.25)+
    geom_sf(data=dat_aged_sf %>% filter(year==yr),col='gray70',size=0.5)+
    geom_sf(data=d1,size=0.5,aes(col=prop))+
    labs(color="Proportion")+
    coord_sf(xlim = c(buff[1],buff[3]),ylim=c(buff[2],buff[4]),expand=FALSE)+
    scale_color_viridis(direction = -1,limits=c(0,1),na.value='gray70')+
    labs(title=paste0(yr,": ",tools::toTitleCase(stg)))+
    theme(legend.position=c(0.7,0.8),legend.background = element_rect(color="white"))
  ggsave(here('model output','maps',stg,paste0('Map ',yr,' ',stg,' proportions.png')),p,w=8,h=12)
})

```
## Hovmoller

Make a hovmoller heat map

```{r}
dat_hov <- dat_spatial_ageprop_binned %>% 
  st_as_sf() %>% 
  st_transform(4326) %>% 
  mutate(lat=st_coordinates(.)[,2],lon=st_coordinates(.)[,1]) %>% 
  st_set_geometry(NULL) %>% 
  mutate(latbin=cut(lat,breaks = c(34.5,seq(36,53,by=0.5),58.5))) %>% 
  group_by(year,latbin,binclass) %>% 
  summarise(totn=sum(totn)/1e6) %>% 
  ungroup() %>% 
  mutate(year=factor(year)) %>%
  complete(year,latbin,binclass,fill = list(totn=0)) %>% 
  group_by(year) %>% 
  mutate(prop=totn/sum(totn)) %>% 
  ungroup()
  
p_hov <- dat_hov %>% 
  ggplot(aes(year,latbin,fill=totn))+
  geom_tile()+
  facet_wrap(~binclass)+
  scale_fill_viridis(option="turbo")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5))+
  labs(x="Year",y="Latitude Range",fill="Abundance\n(millions)")
p_hov

p_hov_rel <- dat_hov %>% 
  ggplot(aes(year,latbin,fill=prop))+
  geom_tile()+
  facet_wrap(~binclass)+
  scale_fill_viridis(option="A")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5))+
  labs(x="Year",y="Latitude Range",fill="Relative Abundance")
p_hov_rel
```

## Center of Gravity

Very coarse latitudinal COG, from the survey

```{r}

dat_cog <- dat_spatial_ageprop %>% 
  st_as_sf() %>% 
  st_transform(4326) %>% 
  mutate(lat=st_coordinates(.)[,2],lon=st_coordinates(.)[,1]) %>% 
  st_set_geometry(NULL) %>% 
  mutate(binclass=ifelse(age%in%c(1,2,3),"juvenile","adult")) %>% 
  group_by(year,binclass) %>% 
  summarise(cog=weighted.mean(lat,n)) %>% 
  ungroup() %>% 
  mutate(year=factor(year))

glimpse(dat_cog)

p_cog <- dat_cog %>% 
  ggplot(aes(year,cog,group=binclass,color=binclass))+
  geom_line()+
  geom_hline(yintercept=44,linetype=2,color="black")+
  geom_hline(yintercept=40,linetype=2,color="gray50")+
  scale_y_continuous(breaks=seq(39,51,by=2))+
  labs(x="Year",y="Mean Latitude",color="Stage")

p_cog
ggsave(here('data','plots','lat_cog_FEAT.png'),p_cog,h=6,w=8)
```


# Animate?

```{r}
list.files(here('model output','maps','juvenile'), pattern = '*.png', full.names = TRUE) %>% 
        image_read() %>% # reads each path file
        image_join() %>% # joins image
        image_animate(delay=300) %>% # animates, can opt for number of loops
        image_write(here('model output','maps','juvenile',"juveniles.gif")) # write to current dir
list.files(here('model output','maps','adult'), pattern = '*.png', full.names = TRUE) %>% 
        image_read() %>% # reads each path file
        image_join() %>% # joins image
        image_animate(delay=300) %>% # animates, can opt for number of loops
        image_write(here('model output','maps','adult',"adults.gif")) # write to current dir
list.files(here('model output','maps'), pattern = '*.png', full.names = TRUE) %>% 
        image_read() %>% # reads each path file
        image_join() %>% # joins image
        image_animate(delay=300) %>% # animates, can opt for number of loops
        image_write(here('model output','maps',"samples_map.gif")) # write to current dir
```



